---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev.args = list(type = "cairo") # for anti-aliasing
)
```

# ggChernoff: draw Chernoff faces in ggplot2

<!-- badges: start -->
[![R-CMD-check](https://github.com/Selbosh/ggChernoff/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Selbosh/ggChernoff/actions/workflows/R-CMD-check.yaml)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ggChernoff)](https://cran.r-project.org/package=ggChernoff)
<!-- badges: end -->

This silly package, `ggChernoff`, introduces a `geom_chernoff` geom for [`ggplot2`](http://ggplot2.org).
This works a bit like [`geom_point`](http://docs.ggplot2.org/current/geom_point.html),
but draws little smiley faces (emoticons) instead of dots.

The Chernoff geom has some unique aesthetics, including `smile`,
which makes your faces smile or frown according to the relative magnitude of your continuous variable.
By default, the mean value will generate a straight face :expressionless: while higher values will make smiles :blush: and lower values will draw frowns :cry:. You can customise this using `scale_smile`.
If `smile` is unmapped to a variable, all faces will be happy by default.

Eyebrows are hidden by default, but you can activate them by mapping something to the `brow` aesthetic.
High values make your faces angry :angry: and low values the opposite :anguished:.

## Installation

You can install the release version of ggChernoff using:

```r
install.packages("ggChernoff")
```

You can install the development version of ggChernoff like so:

``` r
# install.packages("pak")
pak::pak('Selbosh/ggChernoff')
```

## Examples

```{r load}
library(ggplot2)
library(ggChernoff)
```

Firstly, let's create a scatter plot of smiley faces out of Fisher's iris data set, each one coloured according to species.

```{r iris}
ggplot(iris) +
  aes(Petal.Width, Petal.Length, fill = Species) +
  geom_chernoff()
```

Here is an example using Immer's barley data. We are happy about larger yields!

```{r barley}
ggplot(lattice::barley) +
  aes(year, variety, smile = yield, brow = yield) +
  geom_chernoff(fill = 'goldenrod1') +
  scale_x_discrete(limits = c('1931', '1932')) +
  facet_wrap(~site)
```

Basic legends are now supported. We can customise breaks and titles in the usual `ggplot2` way, via `scale_smile_continuous`.

```{r legends}
g <- ggplot(data.frame(x = rnorm(20), y = rexp(20), z = runif(20))) +
  aes(x, y, smile = z) +
  geom_chernoff(fill = 'steelblue1')
g
g + scale_smile_continuous('Smilez', breaks = 0:10 / 10, midpoint = 0.5)
```

You can also use this command to adjust the range of possible happiness/sadness in your plot. In the following example, everybody is somewhere between sad and straight-faced.

```{r range}
g + scale_smile_continuous(range = c(-1, 0))
```

**New feature as of 0.3.0**: eye separation.

```{r eyes}
ggplot(iris) +
  aes(Petal.Width, Petal.Length, fill = Species, eyes = Sepal.Length) +
  geom_chernoff() +
  scale_eyes_continuous(range = c(0, 2))
```

## Space invaders! :alien:

```{r spaceinvaders}
cannon <- data.frame(x = 0, y = 0, colour = 'white', size = 20)

bunkers <- data.frame(x = seq(-4, 4, l = 4), y = 2, colour = 'green')

ufos <- data.frame(
  x = rep(seq(-6, 6, length.out = 12), 5),
  y = rep(6:10, each = 12),
  size = 10,
  colour = c('cyan', 'yellow', 'magenta')[
    c(
      rep(1:3, 4),
      rep(c(2, 3, 1), 4),
      rep(c(3, 1, 2), 4),
      rep(1:3, 4),
      rep(c(2, 3, 1), 4)
    )
  ],
  brow = rnorm(60)
)

ggplot() +
  aes(x, y, fill = colour) +
  geom_chernoff(
    data = ufos,
    aes(size = size, brow = brow),
    smile = -1,
    eyes = 0.5
  ) +
  geom_chernoff(data = cannon, aes(size = size)) +
  geom_tile(data = bunkers, width = 1) +
  geom_tile(
    data = data.frame(x = 0, y = 3, colour = 'white'),
    width = 0.1
  ) +
  scale_fill_identity() +
  scale_size_identity() +
  theme_void() +
  theme(
    plot.background = element_rect(fill = 'black'),
    legend.position = 'none'
  )
```

## References

Hermann Chernoff (1973).
The use of faces to represent points in *k*-dimensional space graphically.
*Journal of the American Statistical Association, 68*(342), 361–368.

Leland Wilkinson (2006).
*The Grammar of Graphics* (2nd edition).
Springer.
